\TheTitle

### 2020-05-19 Paper1/ResponseToReviewers/Response_to_Reviewer.tex

is the i-th moment (scalar function of x and t), while $\bm{\hat u}_j$ in line 225 page
6 is something else (a vector, the index is the cell number). I recommend
to check all these kinds of formal inaccuracies.
\\ \textbf{Response:} We added a time index to $\bm{\hat u}_j$ which was missing in the original version. Furthermore, we added a comment when introducing the moments at the spacial cells in line 189 in order to avoid confusion.

@@

\\ \textbf{Response:} Algorithm 1 is very similar to the original IPM algorithm by Poette. The only difference is that we are not using the old moments $\bm{\hat u}_j^{n+1}$ in the finite volume update (line 11 in algorithm 1), but we are using the term $\langle \bm u_{s}((\bm{\lambda}_j^n)^T\bm{\varphi})\bm{\varphi}^T\rangle_Q^T$. This step ensures that the error from Newton's method does not yield non-physical moments which can lead to a failure of the method, see Section 4.3 in [8].
\item The formulation of Th. 1 is incomplete because $\bm\lambda_*$ is not specified. Are
conditions on $\bm\lambda_*$? It must be explained.
\\ \textbf{Response:} We added an explanation for $\bm\lambda^*$, see line 312. Furthermore, to underline that $\bm\lambda^*$ collects the dual variables at all spatial cells, we now denote $\bm\lambda^*$ as $\bm{\lambda}_{\Delta}^*=(\bm{\lambda}_{1}^*,\cdots,\bm{\lambda}_{N_x}^*)^T$.

@@

notations.
\\ \textbf{Response:} We have changed this in the manuscript.
\item Line 496 page 14, I do not understand the need to specify the temperature $T$ since it does not show up in line 487.
\\ \textbf{Response:} We added a short description on how the defined quantities relate to the conserved variables in line 517.

### 2020-07-03 Paper1/SignificanceAndNovelty/highlights.tex
Paper version elsevier

\subsection*{Highlights}
- intrusive methods for uncertainty quantification have high numerical costs\\
- adaptivity and the One-Shot approach can be used to accelerate intrusive methods\\
- the proposed techniques are studied for an uncertain NACA0012 and pipe testcase with uncertainties up to order three\\
- using the discussed acceleration techniques allows intrusive methods to compete with collocation methods \\
- in our numerical experiments, intrusive methods show a faster convergence to steady state and an improved accuracy compared to collocation methods
\end{document}

>>>>

\subsection*{Highlights}
- intrusive methods for uncertainty quantification have high numerical costs\\
- adaptivity and the One-Shot approach can be used to accelerate intrusive methods\\
- proposed techniques are studied for uncertainties up to order three\\
- acceleration techniques allow intrusive methods to compete with collocation methods
\end{document}

### 2020-07-03 Paper1/framework.tex
Paper version elsevier

\section{Discretization of the IPM system}
\label{sec:framework}
\subsection{Finite Volume Discretization}
In the following, we discretize the moment system in space and time according to \cite{kusch2017maximum}. Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system. Hence, we discretize the system \eqref{eq:SGMomentSystem} with the more general IPM solution ansatz \eqref{eq:ansatz}.  
Omitting initial conditions and assuming a one-dimensional spatial domain, we can write the IPM system  as
\begin{linenomath*}\begin{align*}
\partial_t \bm{\hat u}+\partial_x \bm{F}(\bm{\hat u}) = \bm{0}
\end{align*}\end{linenomath*}
with the flux $\bm{F}:\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$. Note that the inner transpose represents a dyadic product and therefore the outer transpose is applied to a matrix. Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments. We choose the discrete unknowns which represent the solution to be the spatial averages over each cell at time $t_n$, given by
\begin{linenomath*}\begin{align*}
\bm{\hat u}_{ij}^n \simeq \frac{1}{\Delta x}\int_{x_{j-1/ 2}}^{x_{j+1/ 2}}\bm{\hat u}_i(t_n,x) dx.
\end{align*}\end{linenomath*}
If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{ij}^n)_{\vert i\vert\leq M}\in\mathbb{R}^{N\times m}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
\begin{linenomath*}\begin{align}\label{eq:IPMDiscretization}
\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}  - \frac{\Delta t}{\Delta x}\left( \bm{G}(\bm{\hat u}_{j}^{n},\bm{\hat u}_{j+1}^{n})- \bm{G}(\bm{\hat u}_{j-1}^{n},\bm{\hat u}_{j}^{n})\right)
\end{align}\end{linenomath*}
for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$. Here, the number of spatial cells is denoted by $N_x$ and the number of time steps by $N_t$.
The numerical flux is assumed to be consistent, i.e. $\bm{G}(\bm{\hat{u}},\bm{\hat{u}})=\bm{F}(\bm{\hat{u}})$.

@@

\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j).

>>>>

\section{Discretization of the IPM system}
\label{sec:framework}
\subsection{Finite Volume Discretization}
In the following, we discretize the moment system in space and time according to \cite{kusch2017maximum}. Due to the fact, that stochastic-Galerkin can be interpreted as IPM with a quadratic entropy, it suffices to only derive a discretization of the IPM moment system. Hence, we discretize the system \eqref{eq:SGMomentSystem} with the more general IPM solution ansatz \eqref{eq:ansatz}.  
Omitting initial conditions and assuming a one-dimensional spatial domain, we can write the IPM system  as
\begin{linenomath*}\begin{align*}
\partial_t \bm{\hat u}+\partial_x \bm{F}(\bm{\hat u}) = \bm{0}
\end{align*}\end{linenomath*}
with the flux $\bm{F}:\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$, $\bm{F}(\bm{\hat u})=\langle \bm f(\mathcal{U}(\bm{\hat u}))\bm{\varphi}^T \rangle^T$. Note that the inner transpose represents a dyadic product and therefore the outer transpose is applied to a matrix. Due to hyperbolicity of the IPM moment system, one can use a finite-volume method to approximate the time evolution of the IPM moments. We choose the discrete unknowns which represent the solution to be the spatial averages over each cell at time $t_n$, given by
\begin{linenomath*}\begin{align*}
\bm{\hat u}_{ij}^n \simeq \frac{1}{\Delta x}\int_{x_{j-1/ 2}}^{x_{j+1/ 2}}\bm{\hat u}_i(t_n,x) dx.
\end{align*}\end{linenomath*}
If a moment vector in cell $j$ at time $t_n$ is denoted as $\bm{\hat u}_j^n = (\bm{\hat u}_{ij}^n)_{\vert i\vert\leq M}\in\mathbb{R}^{N\times m}$, the finite-volume scheme can be written in conservative form with the numerical flux $\bm{G}:\mathbb{R}^{N\times m}\times\mathbb{R}^{N\times m}\to\mathbb{R}^{N\times m}$ as
\begin{linenomath*}\begin{align}\label{eq:IPMDiscretization}
\bm{\hat u}_{j}^{n+1} = \bm{\hat u}_{j}^{n}  - \frac{\Delta t}{\Delta x}\left( \bm{G}(\bm{\hat u}_{j}^{n},\bm{\hat u}_{j+1}^{n})- \bm{G}(\bm{\hat u}_{j-1}^{n},\bm{\hat u}_{j}^{n})\right)
\end{align}\end{linenomath*}
for $j = 1,\cdots,N_x$ and $n = 0,\cdots,N_t$. Here, the number of spatial cells is denoted by $N_x$ and the number of time steps by $N_t$.
The numerical flux is assumed to be consistent, i.e. $\bm{G}(\bm{\hat{u}},\bm{\hat{u}})=\bm{F}(\bm{\hat{u}})$. The moment vector inside a spatial cell, denoted by $\bm{\hat u}_{j}^{n}$ is not to be confused with the moment of order $i$ in \eqref{eq:SGClosure}, which is a continuous function in time and space. Note that whenever the moment vector includes a time index $n$, we are speaking of a discrete quantity. Furthermore, we use the index $j\in\mathbb{N}$ to denote spatial cells and the multi-index $i\in\mathbb{N}_0^p$ to denote the moment order.

@@

\bm{\lambda}^{(l+1)}_j = \bm{d}(\bm{\lambda}_j^{(l)},\bm{\hat{u}}_j^{n}).

### 2020-07-03 Paper1/one_shot_ipm.tex
Paper version elsevier

\comment{We denote the fixed point of the classical IPM iteration \eqref{eq:momentIteration} by $\bm{\hat u}^*_{\Delta} = (\bm{\hat u}_{1}^*,\cdots,\bm{\hat u}_{N_x}^*)^T$, i.e. we have

@@

}

@@

Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}_{\comment{\Delta}}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by

@@

\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \comment{\bm{\hat u}_j^n}\right) \\
&=- \partial_{\bm{\lambda}}\bm{H}(\comment{\bm\lambda_j^n})^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \comment{\bm{\hat u}_j^n}\right).

@@

\comment{Recall that if $\bm{\hat u}_j^n$ and $\bm{\lambda}_j^n$ belong to the fixed point $(\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)$,} we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \comment{\bm{\hat u}_j^n}$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$. For the block $\partial_{\bm{\hat u}} \bm{d}$, we get 

@@

\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\comment{\bm\lambda_j^n})^{-1},

@@

hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$. \comment{To shorten notation, let us define

@@

Then we have}

@@

\frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}\comment{_j}}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j

@@

With the assumptions from Theorem \ref{th:Contractive}, the One-Shot IPM converges locally. I.e. \comment{if we denote the solution at iteration $n$ as $(\bm{\lambda}_{\Delta}^{(n)},\bm{\hat u}_{\Delta}^{(n)})$}, there exists a $\delta>0$ s.t. for all starting points $(\bm{\lambda}_{\comment{\Delta}}^{(0)},\bm{\hat u}_{\comment{\Delta}}^{(0)})\in B_{\delta}(\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)$ we have

@@

\left\Vert (\bm{\lambda}\comment{_{\Delta}^{(n)}},\bm{\hat u}\comment{_{\Delta}^{(n)}}) - (\bm{\lambda}_{\comment{\Delta}}^*,\bm{\hat u}_{\comment{\Delta}}^*)\right\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.

@@

Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda}_{\comment{\Delta}},\bm{\hat u}_{\comment{\Delta}})$. Hence there exists an open ball with radius $\delta$ around the fixed point in which the eigenvalues remain in the interval $(-1,1)$.

>>>>

We denote the fixed point of the classical IPM iteration \eqref{eq:momentIteration} by $\bm{\hat u}^*_{\Delta} = (\bm{\hat u}_{1}^*,\cdots,\bm{\hat u}_{N_x}^*)^T$, i.e. we have

@@

@@

Since we assume that the classical IPM scheme is contractive at its fixed point, we have $\rho (\nabla_{\bm{\hat u}}\bm{\tilde c}(\bm{\hat u}_{\Delta}^*))<1$ with $\nabla_{\bm{\hat u}}\bm{\tilde c}\in\mathbb{R}^{N\cdot N_x\times N\cdot N_x}$ defined by

@@

\partial_{\bm{\lambda}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) &= \bm{I}_d - \bm{H}(\bm\lambda_j^n)^{-1} \cdot \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}\bm{\varphi}^T \rangle^T - \partial_{\bm{\lambda}}\bm{H}(\bm\lambda_j^n)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}_j^n\right) \\
&=- \partial_{\bm{\lambda}}\bm{H}(\bm\lambda_j^n)^{-1} \cdot \left( \langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T - \bm{\hat u}_j^n\right).

@@

Recall that if $\bm{\hat u}_j^n$ and $\bm{\lambda}_j^n$ belong to the fixed point $(\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)$, we have $\langle \bm{u}_s(\bm{\varphi}^T\bm{\lambda}_j^n)\bm{\varphi}^T \rangle^T = \bm{\hat u}_j^n$, hence one obtains $\partial_{\bm{\lambda}} \bm{d}=\bm{0}$. For the block $\partial_{\bm{\hat u}} \bm{d}$, we get 

@@

\partial_{\bm{\hat u}} \bm{d}(\bm{\lambda}_j^n,\bm{\hat u}_j^n) = \bm{H}(\bm\lambda_j^n)^{-1},

@@

hence $\partial_{\bm{\hat u}} \bm{d}$ is a block diagonal matrix. Let us now look at $\partial_{\bm{\lambda}} \bm{c}$ at a fixed spatial cell $j$. To shorten notation, let us define

@@

Then we have

@@

\frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}}\frac{\partial \bm{d}(\bm{\lambda}_{j-1}^{n},\bm{\hat u}_{j-1}^{n})}{\partial \bm{\hat u}} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}} \bm{H}(\bm\lambda)^{-1} = \frac{\partial \bm{c}_j}{\partial \bm{\lambda}_{\ell}} \langle \nabla\bm{u}_s(\bm{\varphi}^T\bm{\lambda}_{j-1}^n)\bm{\varphi}\bm{\varphi}^T \rangle^{-T} = \partial_{\bm{\hat u}_{\ell}}\bm{\tilde c}_j

@@

With the assumptions from Theorem \ref{th:Contractive}, the One-Shot IPM converges locally. I.e. if we denote the solution at iteration $n$ as $(\bm{\lambda}_{\Delta}^{(n)},\bm{\hat u}_{\Delta}^{(n)})$, there exists a $\delta>0$ s.t. for all starting points $(\bm{\lambda}_{\Delta}^{(0)},\bm{\hat u}_{\Delta}^{(0)})\in B_{\delta}(\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)$ we have

@@

\left\Vert (\bm{\lambda}_{\Delta}^{(n)},\bm{\hat u}_{\Delta}^{(n)}) - (\bm{\lambda}_{\Delta}^*,\bm{\hat u}_{\Delta}^*)\right\Vert \rightarrow 0 \qquad \text{ for } n \rightarrow \infty.

@@

Since the roots of a polynomial vary continuously with its coefficients, the eigenvalues of $\bm{J}$ are continuous w.r.t $(\bm{\lambda}_{\Delta},\bm{\hat u}_{\Delta})$. Hence there exists an open ball with radius $\delta$ around the fixed point in which the eigenvalues remain in the interval $(-1,1)$.
